---
title: "MDSR Exercise 1"
author: "Theodore Dounias"
date: "May 6, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../../"))

library(tidyverse)
library(lubridate)
library(dplyr)

full_fbi <- readLines(file("data-raw\HC 2013 (1)\HC 2013.txt", open = "r"), skipNul = TRUE)

incidents_df <- read.csv("data\hatecrime_incidents_2011to13.csv")


incidents_ex <- incidents_df %>%
  select(2, 3) %>%
  mutate(State_code = substr(Agency_ID,1, 2)) %>%
  select(2,3)

```
##Data Manipulation

If is quite often that, when dealing with data, you might find yourself faced with the daunting task of reading in a particularly disorganized or weirdly formatted dataset. Take for example the following, which is in a vector format containing strings:
```{r, echo = TRUE, cache=TRUE}
print(full_fbi[10:20])
```
    
This is a part of the FBI Uniform Report on hate crimes, in the year 2013. Each line represents either a reporting agency, or a hate crime incident. When a reporting agency reports a hate crime, it is listed under it. Each character, or subsets of the characters, are a specific variable. For example, characters 3-6 represent the state code and state abbreviation of either the agency or the incident. In this exercise, we will look at how we might set ourselves up in a favorable position to be able to tackle data such as this. The data for this exercise is...somewhere.
  
####Question 1
  
Create a function that discerns whether a line is a reporting agency, or an incident, and then divides the two into two other vectors. Some useful functions to remember here are: substr(), if(), for(). Remeber to first think about the internal logic of the function!
  
####Question 2
  
To create the final tidy dataset we need to have a way to extract a variable from each string in our vector. We need to create a function that does this, and returns a dataframe with a single column--the variable we want to add--which we would then bind to our whole dataframe. Write r script that reads 3-4 (State Code), and 5-13 (Agency ID) from the incidents vector you created into separate dataframes, and then merge the, into a new df.

**BONUS:**Can you find a way to create the aforementioned function?
   
####Question 3
   
Now you should have a dataframe with state code and the ID of the agency that reported. The data is already in a much more usable format--it should look like this:
```{r, echo = FALSE, cache = TRUE}
incidents_ex[1:10,]
```
  
We can already make some interesting observations! If you think about it, each line here represents an observation. Each observation is a hate crime incident. This means that we can figure out how many hate crimes hapened in each state! To do this, One way to do this is to use the mutate(), group_by(), and summarize() functions available in the tidyverse. Which state had the highest amount of hate crimes? Can you make an educated guess as to why that is so?
   
**BONUS:**One of the necessary things we should do is figure out how many incidents each reporting agency reported; essentially we need to count how many lines begining with "IR" there are after each line beginning with "BH". To do this, one way is to implement a regressive function. Write a function that is able to return the exact number of incidents reported by each agency.


  
