---
title: "MDSR Exercise 1"
author: "Theodore Dounias"
date: "May 6, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(dplyr)

full_fbi <- readLines(file("C:\\Users\\tdounias\\Downloads\\HC 2013 (1)\\HC 2013.txt", open = "r"), skipNul = TRUE)

```
##Data Manipulation

If is quite often that, when dealing with data, you might find yourself faced with the daunting task of reading in a particularly disorganized or weirdly formatted dataset. Take for example the following, which is in a vector format containing strings:
```{r, include = FALSE}
print(full_fbi[10:20])
```
    
This is a part of the FBI Uniform Report on hate crimes, in the year 2013. Each line represents either a reporting agency, or a hate crime incident. When a reporting agency reports a hate crime, it is listed under it. Each character, or subsets of the characters, are a specific variable. For example, characters 3-6 represent the state code and state abbreviation of either the agency or the incident. In this exercise, we will look at how we might set ourselves up in a favorable position to be able to tackle data such as this.
  
*Question 1*
  
